%%%% Time-stamp: <2012-08-20 17:41:39 vk>

%% example text content
%% scrartcl and scrreprt starts with section, subsection, subsubsection, ...
%% scrbook starts with part (optional), chapter, section, ...

\chapter{Project Specific Implementation}
\label{sec:projectspecific} 

The preceding implementation descriptions focus exclusively on the OBDMiddleware package and thus on functionalities shared
by both, OBD Tool and OBDCU. The resulting OBDMiddleware package is an API.
Furthermore a concrete application has to be implemented to apply the designed features.
The OBDCU, which stands for On Board Diagnosis Control Unit, represents the emulated diagnosis CU.
It offers possibilities to emulate attached USB devices on Linux systems and is able to set and unset
permanent and pending diagnostic trouble codes as well as deleting them.

The OBD tool is supposed to be its counterpart for reading trouble codes and its corresponding freeze frame data.
The OBDCU’s sensor capabilities as well as the OBD tool itself are not implemented due to underestimation of the needed time for unit test 
implementation and research. 
Although it is safe to say that the general functionalities for requesting sensor data is included in the OBDMiddleware. 
Even though the base function set is asserted by unit tests,
the additional implementation and connection of the related sensor functions is beyond the scope of this bachelor thesis.

By using CMake simple compilation and dependency requirements can be ensured.
The latter is particularly important, since dependencies on external and internal packages would be difficult to check otherwise.
A setup for development is provided by making paths to the subprojects configurable. 
Packaging for installation is realized by using CPack generators, especially for Debian systems.

The Model-View-Controller architecture is chosen due to its clear specification of classes and their clustering into general divisions. 
The model being represented by the middleware package itself.

\section{OBDCU}

The OBDCU application offers the user the possibility to emulate and partially configure an USB FTDI device with one configuration
and one set of bulk endpoints. Furthermore diagnostic trouble codes are selectable from a database and settable to the emulated 
fault memory either as a permanent or temporary trouble code. 
Additionally it provides a surface to configure responses to ELM commands, where atz.

The current beta version has several restrictions regarding the configurations. They are only kept in memory while the application is running. 
However it is capable of simulating at least three of ten service modes as well as all ELM commands.
As previously mentioned service modes one and two, which represent reading of sensor and freeze frame data, are prepared in the OBDMiddleware package
but since the project has already exceeded the projected time for this bachelor thesis the implementation was skipped.

\section{Installation How-To}

Several steps have to be carried out before starting or building the project on a development environment.
This project has a lot of dependencies, which are listed in the built package, and it requires kernel modules.
This section elaborates on additional dependencies and building instructions depending on the purpose.

The project uses different open source APIs to realize the desired functionalities.
The following packages are included in the official Ubuntu repositories and can be downloaded by any package management system:

\vbox{
\begin{itemize}
 \item libmysqlcppconn-dev 
 \item mysql-server
 \item libusb 
 \item libusb-dev
 \item libxml++2.6-2 
 \item libxml++2.6-2-dev 
 \item qt-sdk
 \item libcppunit-dev 
 \item libcppunit-1.13-0 
 \item libboost-all-dev 
\end{itemize}
}

The only packages which require manual compilation are the libusb-vhci library and the vhci\_hcd package, 
resulting in two kernel module files (.ko). Those files either have to be inserted manually using insmod 
or they have to be installed with the help of ``make install''. When installing and frequently accessing the 
kernel modules it is advisable to alter the modules system file, located under “/etc/modules”, by adding 
them at the end of the file. This will result in the loading of the kernel modules at startup.

To compile the libusb-vhci project the following steps should be taken:

\paragraph{vhci\_hcd}

\begin{enumerate}
 \item download or use from repository
 \item make
 \item sudo make install
\end{enumerate}

Optional - depending on how kernel modules want to be handled:

\begin{enumerate}[resume]
 \item insmod usb-vhci-hcd.ko
 \item insmod usb-vhci-iocifc.ko
\end{enumerate}
 
or:
 
\begin{enumerate}
\setcounter{enumi}{3}
 \item sudo gedit /etc/modules \&
 \item add kernel module names (without extension) to file and save
\end{enumerate}

\paragraph{libusb\_vhci [userspace tools]}

\begin{enumerate}
 \item download or use from repository
 \item ./configure
 \item make
 \item sudo make install
 \item sudo ldconfig -v 
\end{enumerate}

Depending on the Linux kernel version and the current vhci\_hcd and libusb\_vhci versions, 
the compiling procedure may vary. These instructions are extracted from the INSTALL 
files from the repositories. The last installation step is to tell the Linux ld path that 
libusb-vhci has been added to the library path so that it can be included normally.

\section{Building the Project}

When building the OBD project it is essential that it is built in the right order. 
The OBDMiddleware package has to be built before the project specific parts. 
Each part of the OBDMiddleware package can be compiled separately, although it is not advisable, 
since the packages have internal dependencies. 
With the support of CMake building in any location is possible with the command:

\begin{verbatim}
 cmake <path to parent CMakeLists>
\end{verbatim}

Furthermore a package can be built using make package. This will result in a .deb file for Ubuntu distributions. 
Executing the generated file leads to the installation of the files into the library path of the Linux system. 

Like in most software projects it was deemed necessary, that the project can be built with just a compiled version 
of the middleware in case some changes emerge. Thus the project specific parts can be built without parameters 
if the OBDMiddleware package is installed. When intending to use the compiled version it can be built by delivering 
the CMake variable Middleware\_BASE\_PATH.

\begin{verbatim}
 cmake -DMiddleware_BASE_PATH=/home/<...>/OBDMiddleware/build/ .. 
\end{verbatim}

If the path has changed it is advisable to clear the build directory or at least delete the generated CMakeCache.txt 
since CMake has caching issues and does not reevaluate the given variables but rather uses the stored paths. 



%% vim:foldmethod=expr
%% vim:fde=getline(v\:lnum)=~'^%%%%\ .\\+'?'>1'\:'='
%%% Local Variables: 
%%% mode: latex
%%% mode: auto-fill
%%% mode: flyspell
%%% eval: (ispell-change-dictionary "en_US")
%%% TeX-master: "main"
%%% End: 
